Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
tweak
grammar check

\todo[inline]{
	Check grammar;
	Improve the proof of \ref{thm:SD-bisim}.
	And the proof of \ref{thm:SD-equiv}
}

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{flanagan2006hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
iteroperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses
\citep{siek2015blame,siek2010threesomes,garcia2013calculating,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}.
The current state of the art includes

\begin{itemize}
\item threesomes \citep{siek2010threesomes,garcia2013calculating},
\item supercoercions \citep{garcia2013calculating}, and
\item coercions in normal form
  \citep{siek2012interpretations,siek2015blame}.
\end{itemize}
%Recall that in these systems, casts are compressed using a composition
%operator.
These systems compress casts using a compose operator.
Threesomes and supercoercions are good for mechanized
metatheory because their compose operators are structurally recursive,
making them easy to define in a proof assistant such as Agda. In
contrast, the coercions in normal form have compose operators that are
not structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labelled bottom type), and
supercoercions (10 different kinds).

This paper presents a new cast representation, named
\emph{hypercoercions}, that is good for both mechanized metatheory and
good for implementation. The composition operator for hypercoercions
is defined by structural recursion and hypercoercions are suggestive
of a bit-level representation that minimizes the need for pointers and
fits all first-order casts into 64 bits.
%
We present two flavors of hypercoercions to support the two blame
tracking strategies from the literature: D and
UD~\citep{siek2009exploring}. We are interested in the D blame
tracking strategy because it comes with a more straightforward notion of
safe cast compared to UD~\citep{siek2009exploring}, which is why D was
chosen in Grift compiler~\citep{kuhlenschmidt2018efficient}. We are also 
interested in UD because it plays a
prominent role in the gradual typing literature
\citep{wadler2009well}.  The UD hypercoercions were inspired by the
supercoercions of \citet{garcia2013calculating} (hence the name) and
the D hypercoercions were inspired by the normal forms of
\citet{siek2012interpretations}.
%
The semantics of casts can be lazy or
eager~\citep{siek2009exploring}. In this paper we focus on lazy cast
strategies because we suspect that they are more efficient than eager
strategies and because \citet{new2019gradual} show that eager
strategies are incompatible with $\eta$-equivalence of functions.

@@

To summarize, the primary contributions of this paper are:
\begin{itemize}
\item hypercoercions, a new cast representation, which
  has a structurally recursive composition and a more compact memory
  representation.
\item a framework in Agda for proving the correctness of \lazyD{} cast
  represantations.
\item a formal proof that \lazyD\ hypercoercions respect the semantics
  of the \lazyD{} cast calculus.
\end{itemize}

In Section~\ref{sec:background} we review cast calculi and coercions. 
We present hypercoercions in
Section~\ref{sec:hypercoercion-definition}.  We present a framework
for proving correctness of \lazyD\ cast reprensentations in
Section~\ref{sec:framework} and use it to prove the correctness of
\lazyD{} hypercoercions in
Section~\ref{sec:hypercoercion-correctness}.

@@

The syntax and static semantics is the same for the \lazyD{} and
\lazyUD{} cast calculi, and is reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

The dynamic semantics of a cast calculus is typically defined with a
reduction semantics. Here we use a CEK machine
\citep{felleisen1986control} instead because the first author is more
familiar with CEK machines and believes that a CEK machine is more
convenient to use for the space-efficient semantics in
Section~\ref{sec:framework:cek} (a major point of space-efficiency is about 
compressing continuations). So using an abstract machine for the cast calculi in
this section makes it easier to prove correctness of the space-efficient
machines. Of course, one should prove that the abstract machine
presented here is equivalent to the standard reduction semantics for
cast calculi, but we have not yet done so.

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split a casted injection, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

Next we define the CEK Machine for \lazyUD{}. The only difference with
respect to \lazyD{} is in the definition of the $\mathit{applyCast}$
function, in which a cast whose source or target is the unknown type
$\star$ is always split into two casts that go through an injectiable
type, that is, a type in which all sub-components are the unknown
type, such as $\star \to \star$.

@@

In this section we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal
form of \citet{siek2012interpretations} to motivate the design of
hypercoercions.  We omit sum types and product types in this section,
because \citet{siek2012interpretations} did not discuss them. We
assume a basic familiarity with coercions and direct readers who are
not familiar with them to \citet{siek2012interpretations} and
\citet{siek2015blame}.

@@

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $I!$ takes a value from an injectiable type
$I$ to type $\TOOdyn$. An injectable type is simply a type that can be
cast directly to (injection) and from (projection) $\TOOdyn$. The
definition of injectable type depends on blame strategy. With \lazyD,
all pre-types are injectable. With \lazyUD, only
$\POOfun{\TOOdyn}{\TOOdyn}$ is injectable.
%
A projection $I?^l$ takes a value from type $\TOOdyn$ to type $I$, or
halts the program and blames $l$ if the value is of a different type.
The coercion $\iota$ is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\ncSeq{c_1}{c_2})$ applies $c_1$ and
then $c_2$ in sequence. The function coercion $\ncFun{c_1}{c_2}$
applies $c_1$ to the argument of a function and $c_2$ to the return
value.

@@

This section presents our first contribution, the definition of
hypercoercions.  The design of hypercoercions is motivated by the
observation that a normal coercion has at most three
parts. Hypercoercions make this structure explicit: a hypercion $c$ is
either $\hyperCoercionI$, the identity cast for $\TOOdyn$, or it contains
three parts: a head, middle, and tail, as defined in
\figref{fig:hypercoercion}.
\begin{itemize}
\item The head $h$ is either a projection or the no-op,
\item the middle $m$ involves coercions that preserve a type
  constructor, i.e., coercions between function, pair, sum types, and
  unit types, and
\item the tail $t$ is either an injection, a failure, or the no-op.
\end{itemize}
There is a close correspondence between middle coercions and type
constructors.  We generalize shallow-consistency to middles $m \smile
m$ in the obvious way.

Subsection~\ref{sec:ld-hc} defines the $id$, $seq$, and $cast$
functions that construct
\lazyD\ hypercoercions. Subsection~\ref{sec:lud-hc} defines the
\lazyUD\ counterparts.  The functions that apply hypercoercions to
values are defined later in
Section~\ref{sec:hypercoercion-correctness}, after we introduce the framework 
and its new space-efficient definition of values in Section~\ref{sec:framework}.

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]

  Build continuation \fbox{$wrap(k) = \kappa$}
  \[
  \begin{array}{rclc}
  \funrule{wrap(k)}{\kOOcast{id(T_1)}{k}}{
    \sidecond{k : T_1 \Longrightarrow T_2}}
  \end{array}
  \]

@@

Before proving the weak bisimulation between \effCEK{C} and \ineffCEKD, we 
introduce one more definition. $s_1 \approx^{*} s_2$ means we will get $s_1 
\approx s_2$ after more transition steps in \ineffCEKD.

\begin{definition}[$s \approx^{*} s$]
Given any two states $s_1$ and $s_2$, we write $s_1 \approx^{*} s_2$
if and only if there exists a state $s_2' \in \text{\ineffCEKD}$ such
that
\[
\judgeDreduceTrans{s_2}{s_2'}
\;\text{and}\;
s_1 \approx s_2'
\]
\end{definition}

%\todo[inline]{Explain that this is an unusual formulation of a weak
%  bisimulation.  We promise it will all be OK regarding how it is used
%  in the correctness proof for S(C). -JS}

The lemma below formulate the weak bisimulation between \effCEK{C} and 
\ineffCEKD. It is unusual because we use $s \approx s$ in premise while $s 
\approx^{*} s$ in conclusion. As we will show in the correctness proof for 
\effCEK{C}, however, this formulation is fine for our purpose.

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
{\color{red}
  Jeremy's attempt to reformulate in a more standard way that
  remains equivalent:
\begin{enumerate}
\item If $\judgeSreduce{C}{s_1}{s_3}$ then
  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
\item If $\judgeDreduce{s_2}{s_4}$ then
  $\judgeSreduce{C}{s_1}{s_3}$,
  $\judgeDreduceTrans{s_4}{s_5}$, and
   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
}
Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follow from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

	The other direction is similar.
%  Proving that the left-hand side implies the right-hand side is
%  trivial.  To prove $\judgeDeval{e}{o}$ implies
%  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
%  catch it up in \ineffCEKD\ by taking one or more steps in
%  \ineffCEKD. As we just took a positive number of steps, we are
%  closer to the halting state.
\end{proof}

>>>>

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{flanagan2006hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
interoperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses
\citep{siek2015blame,siek2010threesomes,garcia2013calculating,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}.
The current state of the art includes

\begin{itemize}
\item threesomes \citep{siek2010threesomes,garcia2013calculating},
\item supercoercions \citep{garcia2013calculating}, and
\item coercions in normal form
  \citep{siek2012interpretations,siek2015blame}.
\end{itemize}
%Recall that in these systems, casts are compressed using a composition
%operator.
These systems compress casts using a compose operator.
Threesomes and supercoercions are good for mechanized
metatheory because their compose operators are structurally recursive,
making them easy to define in a proof assistant such as Agda. In
contrast, the coercions in normal form have compose operators that are
not structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labeled bottom type), and
supercoercions (10 different kinds).

This paper presents a new cast representation, named
\emph{hypercoercions}, that is good for both mechanized metatheory and
good for implementation. The composition operator for hypercoercions
is defined by structural recursion and hypercoercions are suggestive
of a bit-level representation that minimizes the need for pointers and
fits all first-order casts into 64 bits.
%
We present two flavors of hypercoercions to support the two blame
tracking strategies from the literature: D and
UD~\citep{siek2009exploring}. We are interested in the D blame
tracking strategy because it comes with a more straightforward notion of
safe cast compared to UD~\citep{siek2009exploring}, which is why D was
chosen in Grift compiler~\citep{kuhlenschmidt2018efficient}. We are also 
interested in UD because it plays a
prominent role in the gradual typing literature
\citep{wadler2009well}.  The UD hypercoercions were inspired by the
supercoercions of \citet{garcia2013calculating} (hence the name) and
the D hypercoercions were inspired by the normal forms of
\citet{siek2012interpretations}.
%
The semantics of casts can be lazy or
eager~\citep{siek2009exploring}. In this paper, we focus on lazy cast
strategies because we suspect that they are more efficient than eager
strategies and because \citet{new2019gradual} show that the eager
strategies are incompatible with $\eta$-equivalence of functions.

@@

To summarize, the primary contributions of this paper are:
\begin{itemize}
\item hypercoercions, a new cast representation, which
  has a structurally recursive composition and a more compact memory
  representation.
\item a framework in Agda for proving the correctness of \lazyD{} cast
  representations.
\item a formal proof that \lazyD\ hypercoercions respect the semantics
  of the \lazyD{} cast calculus.
\end{itemize}

In Section~\ref{sec:background} we review cast calculi and coercions. 
We present hypercoercions in
Section~\ref{sec:hypercoercion-definition}.  We present a framework
for proving the correctness of \lazyD\ cast representations in
Section~\ref{sec:framework} and use it to prove the correctness of
\lazyD{} hypercoercions in
Section~\ref{sec:hypercoercion-correctness}.

@@

The syntax and static semantics is the same for the \lazyD{} and
\lazyUD{} cast calculi and is reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

The dynamic semantics of a cast calculus is typically defined with a
reduction semantics. Here we use a CEK machine
\citep{felleisen1986control} instead because the first author is more
familiar with CEK machines and believes that a CEK machine is more
convenient to use for the space-efficient semantics in
Section~\ref{sec:framework:cek} (a major point of space-efficiency is about 
compressing continuations). So using an abstract machine for the cast calculi in
this section makes it easier to prove the correctness of the space-efficient
machines. Of course, one should prove that the abstract machine
presented here is equivalent to the standard reduction semantics for
cast calculi, but we have not yet done so.

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or continuing $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly takes out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split a casted injection, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

Next, we define the CEK Machine for \lazyUD{}. The only difference with
respect to \lazyD{} is in the definition of the $\mathit{applyCast}$
function, in which a cast whose source or target is the unknown type
$\star$ is always split into two casts that go through an injectable
type, that is, a type in which all sub-components are the unknown
type, such as $\star \to \star$.

@@

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal
form of \citet{siek2012interpretations} to motivate the design of
hypercoercions.  We omit sum types and product types in this section
because \citet{siek2012interpretations} did not discuss them. We
assume a basic familiarity with coercions and direct readers who are
not familiar with them to \citet{siek2012interpretations} and
\citet{siek2015blame}.

@@

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $I!$ takes a value from an injectable type
$I$ to type $\TOOdyn$. An injectable type is simply a type that can be
cast directly to (injection) and from (projection) $\TOOdyn$. The
definition of injectable types depends on the blame strategy. With \lazyD,
all pre-types are injectable. With \lazyUD, only
$\POOfun{\TOOdyn}{\TOOdyn}$ is injectable.
%
A projection $I?^l$ takes a value from type $\TOOdyn$ to type $I$, or
halts the program and blames $l$ if the value is of a different type.
The coercion $\iota$ is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\ncSeq{c_1}{c_2})$ applies $c_1$ and
then $c_2$ in sequence. The function coercion $\ncFun{c_1}{c_2}$
applies $c_1$ to the argument of a function and $c_2$ to the return
value.

@@

This section presents our first contribution, the definition of
hypercoercions.  The design of hypercoercions is motivated by the
observation that a normal coercion has at most three
parts. Hypercoercions make this structure explicit: a hypercoercion $c$ is
either $\hyperCoercionI$, the identity cast for $\TOOdyn$, or it contains
three parts: a head, middle, and tail, as defined in
\figref{fig:hypercoercion}.
\begin{itemize}
\item The head $h$ is either a projection or the no-op,
\item the middle $m$ involves coercions that preserve a type
  constructor, i.e., coercions between function, pair, sum types, and
  unit types, and
\item the tail $t$ is either an injection, a failure, or the no-op.
\end{itemize}
There is a close correspondence between middle coercions and type
constructors.  We generalize shallow-consistency to middles $m \smile
m$ in the obvious way.

Subsection~\ref{sec:ld-hc} defines the $id$, $seq$, and $cast$
functions that construct
\lazyD\ hypercoercions. Subsection~\ref{sec:lud-hc} defines
\lazyUD\ counterparts.  The functions that apply hypercoercions to
values are defined later in
Section~\ref{sec:hypercoercion-correctness}, after we introduce the framework 
and its new space-efficient definition of values in Section~\ref{sec:framework}.

@@

\begin{definition}[Continuations for the \effCEK{C} machine] 
\label{def:values-effCEK}
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]
\end{definition}

\begin{figure} 
  Build continuation \fbox{$wrap(k) = \kappa$}
  \[
  \begin{array}{rclc}
  \funrule{wrap(k)}{\kOOcast{id(T_1)}{k}}{
    \sidecond{k : T_1 \Longrightarrow T_2}}
  \end{array}
  \]

@@

%Before proving the weak bisimulation between \effCEK{C} and \ineffCEKD, we 
%introduce one more definition. $s_1 \approx^{*} s_2$ means we will get $s_1 
%\approx s_2$ after more transition steps in \ineffCEKD.
%
%\begin{definition}[$s \approx^{*} s$]
%Given any two states $s_1$ and $s_2$, we write $s_1 \approx^{*} s_2$
%if and only if there exists a state $s_2' \in \text{\ineffCEKD}$ such
%that
%\[
%\judgeDreduceTrans{s_2}{s_2'}
%\;\text{and}\;
%s_1 \approx s_2'
%\]
%\end{definition}
%
%%\todo[inline]{Explain that this is an unusual formulation of a weak
%%  bisimulation.  We promise it will all be OK regarding how it is used
%%  in the correctness proof for S(C). -JS}
%
%The lemma below formulate the weak bisimulation between \effCEK{C} and 
%\ineffCEKD. It is unusual because we use $s \approx s$ in premise while $s 
%\approx^{*} s$ in conclusion. As we will show in the correctness proof for 
%\effCEK{C}, however, this formulation is fine for our purpose.

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $\judgeSreduce{C}{s_1}{s_3}$ then
	$\judgeDreduceTrans{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
	\item If $\judgeDreduce{s_2}{s_4}$ then
	$\judgeSreduceTrans{C}{s_1}{s_3}$,
	$\judgeDreduceTrans{s_4}{s_5}$, and
	$s_3 \approx s_5$ for some $s_3$ and $s_5$.
	\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
%\begin{enumerate}
%  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
%  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
%    then $\judgeDreduce{s_2}{s_4}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_4$
%  \item If $\judgeDreduce{s_2}{s_4}$
%    then $\judgeSreduce{C}{s_1}{s_3}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_3$
%\end{enumerate}
%{\color{red}
%  Jeremy's attempt to reformulate in a more standard way that
%  remains equivalent:
%\begin{enumerate}
%\item If $\judgeSreduce{C}{s_1}{s_3}$ then
%  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
%\item If $\judgeDreduce{s_2}{s_4}$ then
%  $\judgeSreduce{C}{s_1}{s_3}$,
%  $\judgeDreduceTrans{s_4}{s_5}$, and
%   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
%\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%\end{enumerate}
%}
%Or diagrammatically, 
%  \[\begin{array}{clclc}
%  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%  \rotatebox[origin=c]{90}{$\approx$} 
%  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%  s_{2} & 
%  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
%  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follow from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

	Now let us prove right to left by induction on 
	$\judgeDreduceTrans{s_2}{\sOOhalt{o}}$.
	\begin{itemize}
		\item If $s_2$ takes zero step to the halting state, it must be halting.
		By Lemma~\ref{thm:SD-bisim}, $s_2$ is halting as well, so we are done.
		\item If $\judgeDreduce{s_2}{s_4}$ and 
		$\judgeDreduceTrans{s_4}{\sOOhalt{o}}$,
		by Lemma~\ref{thm:SD-bisim}, there exist $s_3,s_5$ such that
		\[
			\judgeSreduceTrans{C}{s_1}{s_3}
			\;\text{and}\;
			\judgeDreduceTrans{s_4}{s_5}
			\;\text{and}\;
			s_3 \approx s_5
		\]
		From $\judgeDreduceTrans{s_4}{\sOOhalt{o}}$ and 
		$\judgeDreduceTrans{s_4}{s_5}$ we know 
		$\judgeDreduceTrans{s_5}{\sOOhalt{o}}$ because our machine is 
		deterministic. Now we can apply our induction hypothesis.
	\end{itemize}
%  Proving that the left-hand side implies the right-hand side is
%  trivial.  To prove $\judgeDeval{e}{o}$ implies
%  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
%  catch it up in \ineffCEKD\ by taking one or more steps in
%  \ineffCEKD. As we just took a positive number of steps, we are
%  closer to the halting state.
\end{proof}
